MapReduce to My Ears\\ {\Large Using Big Data for Music Synthesis}

### 2013-05-10 writeup/paper.tex
adding paper

\usepackage{listings}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{amssymb}
\usepackage{url}
\usepackage[splitrule]{footmisc}

@@

\title{Using Big Data for Music Synthesis}

@@

The core idea in converting image data to music is to view the
information in an image (color, size, opacity, etc..) as a mapping to
notes and durations with which those notes are sustained.  In our
system we chose MIDI as the output format.  Our system can then be
modeled as a function \kris{transducer?} from image data to MIDI
output.  Note that the choice of MIDI limits the expressivity of our
music.  Going to other formats (e.g., MP3, OOG, etc..) could provide a
larger range of primitives with which to construct music, but does not
enjoy the simplicity that MIDI provides.

Our system is modeled as a function from ordered sets of images to a
MIDI file.  In our view, a MIDI file consists of note and duration
pairs.  Taken together, these constitute a score by chaining together
multiple pairs over the course of time:

@@

%% \[
%%     f : image^{\ast} \to (
%% \]

@@

To convert image data to musical notes, we need to establish a
bijection between the data in the image and the range of music in the
song.  The way we do this is inspired by viewing colors in images as
HSV values.  In HSV, each color has three components: a hue, a
saturation, and a value.  These components are visualized (as shown in
Figure~\ref{fig:hsv}) using a wheel.  The components represent:

\begin{enumerate}
\item Hue --- The actual ``color'' that would be described.  At full
  saturation and value, the hue determines the color seen.  This is
  visualized as going around a wheel of colors.

\item Saturation --- The degree to which the color is saturated from
  white.  Zero saturation is a blank (white) surface, while increasing
  values show more permeance of color.

\item Value --- The amount of light in the color.  Colors with low
  values are dark, and those with higher values are brighter.  This is
  visualized by going up the cylinder.
\end{enumerate}

\kris{Here I'm going to draw a diagram which shows two points on the
  HSV wheel and a musical clef, illustrating the position.}

In our transformation to music, we map colors in HSV to notes in MIDI.
To do this, we map colors in RGB to HSV using the \code{\#RGBtoHSV}
method of the \code{java.awt.Color} class.

\begin{figure}
  \centering
  \includegraphics[width=.4\textwidth]{hsv.png}
  \caption{The HSV color cylinder}
  \label{fig:hsv}
\end{figure}

@@

\begin{figure}
\centering
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{image1}
  \caption{A subfigure}
  \label{fig:sub1}
\end{subfigure}%
\begin{subfigure}{.5\textwidth}
  \centering
  \includegraphics[width=.4\linewidth]{image1}
  \caption{A subfigure}
  \label{fig:sub2}
\end{subfigure}
\caption{A figure with two subfigures}
\label{fig:test}
\end{figure}

@@

Our implementation consists of scripts to transform the input,
MapReduce phases, and a music mixing tool.  The high level pipeline is
shown in Figure~\ref{fig:project-pipeline}.

\paragraph{Conversion of images to thumbnails}

@@

\paragraph{Conversion to intermediate representation}
\label{sec:conversion}

@@

\begin{figure}
\begin{lstlisting}[language=Java]
public class WriteImagesToSequenceFile {
  BufferedImage img = null;
  SequenceFile.Writer writer = null;
  FileSystem fs = FileSystem.get(conf);

  FileStatus[] fss = 
    fs.globStatus(new Path(inputPath 
      + "/*.jpg"));

  writer = SequenceFile.createWriter(fs, 
    conf, outPath, IntWritable.class,
    ArrayListOfIntsWritable.class);
  int counter = 0;
  for (FileStatus status : fss) {
    IntWritable image_counter = 
      new IntWritable();
    ArrayListOfIntsWritable 
      pixelData = 
        new ArrayListOfIntsWritable();

    img = ImageIO.read(
      new File(status.getPath()
               .toString().substring(5)));

    // Get the RBG integer for each pixel 
    // in 3X3
    for(int j = 0; j < thumbs; j++) {
      for(int i = 0; i< thumbs; i++) {
        pixelData.add(img.getRGB(i,j));
      }
    }

    // Write to the sequence file
    image_counter.set(counter);
    writer.append(image_counter, pixelData);
    counter += 1;
  } 
}
\end{lstlisting}
\caption{Snippet of code from \code{WriteToSequenceFile} showing
  format used in sequence files throughout the MapReduce jobs.}
\end{figure}

\paragraph{Translating to note, region pairs}

@@

The strategy used in the conversion is to map over each pixel in an
image, apply the conversion function, and emit it to a reducer to
collate together.  After mapping over each pixel, we need to end up
with a stream of images for each region: a naive implementation simply
maps each pixel in our input corpus to these tone, velocity pairs:

@@

>>>>

\usepackage{listings}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{rotating}
\usepackage{geometry}
\usepackage{amssymb}
\usepackage{url}
\usepackage[splitrule]{footmisc}

@@

\title{MapReduce to My Ears\\
{\Large Using Big Data for Music Synthesis}
}

@@

The core idea in converting image data to music is to view the
information in an image (color, size, opacity, etc..) as a mapping to
notes and durations with which those notes are sustained.  In our
system we chose MIDI as the output format.  Our system can then be
modeled as a function from image data to MIDI output.  Note that the
choice of MIDI limits the expressivity of our music.  Going to other
formats (such as MP3 or OOG) could provide a larger range of
primitives with which to construct music, but does not enjoy the
simplicity that MIDI provides.

Our system takes in ordered sets of images and produces a MIDI file.
In our view, a MIDI file consists of note and duration pairs.  Taken
together, these constitute a score by chaining together multiple pairs
over the course of time:

@@

However, producing a MIDI score with only a single instrument would
sound broing (akin to a song with only one player): instead we take
pieces of images and map them into different parts.  Each of these
parts constitutes an instrument and --- when mixed together ---
produces a final score.  An illustration of the core idea is shown in
Figure~\ref{fig:images-to-music}.

\begin{figure*}
  \centering
  \begin{minipage}[t]{.3\textwidth}
    \centering
    \includegraphics[width=.7\textwidth]{images-to-regions.png}
  \end{minipage}
  \begin{minipage}[t]{.1\textwidth}
    \raisebox{.8 in}{\Huge $\to$}
  \end{minipage}
  \begin{minipage}[t]{.4\textwidth}
    \centering
    \includegraphics[width=\textwidth]{parts-and-instruments.png}
  \end{minipage}
  \caption{Mapping a stream of three images to a nine piece musical
    score.  First each image is mapped into regions, which will be
    transformed into notes and combined into a score using a mixer.}
\end{figure*}

\begin{figure}[h]
  \centering
  \includegraphics[width=.4\textwidth]{hsv.png}
  \caption{The HSV color cylinder}
  \label{fig:hsv}
\end{figure}

@@

To convert image data to musical notes, we need to establish a
bijection between the data in the image and the range of music in the
song.  The way we do this is inspired by viewing colors in images as
HSV values.  In HSV, each color has three components: a hue, a
saturation, and a value.  These components are visualized (as shown in
Figure~\ref{fig:hsv}) as a cylinder (as opposed to RGB, which views
the color space as a cube).  Each component represents:

\begin{enumerate}
\item Hue: The actual ``color'' that would be described.  At full
  saturation and value, the hue determines the color seen.  This is
  visualized as going around a wheel of colors.

\item Saturation: The degree to which the color is saturated from
  white.  Zero saturation is a blank (white) surface, while increasing
  values show more permeance of color.

\item Value: The amount of light in the color.  Colors with low values
  are dark, and those with higher values are brighter.  This is
  visualized by going up the cylinder.
\end{enumerate}

In our transformation to music, we map colors in HSV to notes in MIDI.
We do this by overlaying the circle of fifths (a circle representing
pleasing intervals called fifths in music theory) onto the circle
representing hue in HSV.  We establish the octave of the produced note
using the value of the color in HSV, and the volume of the note using
the saturation.

@@

@@

Our implementation consists of scripts to transform the input,
MapReduce phases, and a music mixing tool.  The high level pipeline is
shown in Figure~\ref{fig:project-pipeline}.  It consists of a sequence
of MapReduce jobs, along with sets of helper scripts

\subsection{Conversion of images to thumbnails}

@@

\subsection{Conversion to intermediate representation}
\label{sec:conversion}

@@

\begin{figure}
\begin{lstlisting}[language=Java]
public class WriteImagesToSequenceFile
{
  BufferedImage img = null;
  SequenceFile.Writer writer = null;
  FileSystem fs = 
      FileSystem.get(conf);
  FileStatus[] fss = 
    fs.globStatus(new Path(input 
      + "/*.jpg"));
  writer = SequenceFile
      .createWriter(fs, 
    conf,outPath,IntWritable.class,
    ArrayListOfIntsWritable.class);
  int counter = 0;
  for (FileStatus status : fss) {
    IntWritable image_counter = 
      new IntWritable();
    ArrayListOfIntsWritable 
      pixelData = 
        new ArrayListOfIntsWritable();
    img = ImageIO.read(
      new File(status.getPath()
           .toString().substring(5)));
    // Get RBG integer for each pixel 
    for(int j = 0; j < thumbs; j++) {
      for(int i = 0; i< thumbs; i++) {
        pixelData
          .add(img.getRGB(i,j));
      }
    }
    // Write to the sequence file
    image_counter.set(counter);
    writer.append(image_counter,
      pixelData);
    counter += 1;
  } 
}
\end{lstlisting}
\caption{Snippet of code from \code{WriteToSequenceFile} showing
  format used in sequence files throughout the MapReduce jobs.}
\end{figure}

\subsection{Translating to note, region pairs}

@@

The strategy used in the conversion is to map over each pixel in an
image, apply the conversion function, and emit it to a reducer to
collate together.  After mapping over each pixel, we need to end up
with a stream of images for each region: a na\"{i}ve implementation simply
maps each pixel in our input corpus to these tone, velocity pairs:

@@

A snippet of our mapper is shown in
Figure~\ref{figure:colortomusic-mapper}.

\begin{figure}
\label{figure:colortomusic-mapper}
\begin{lstlisting}[language=Java]
public void map(IntWritable key, 
    ArrayListOfIntsWritable pixels,
    Context context)
  throws IOException, InterruptedException {
  int regionCounter = 1;
  MidiNote midiNote = null;
  for(int i = 0; i < pixels.size(); i++){
    pixel = pixels.get(i);
    // Conversion from color to note
    midiNote = Color2Music
      .convert(new Color(pixel));
    REGION_IMAGENO.set(regionCounter,
      key.get());
    NOTE_VELOCITY.set(midiNote.getTone(),
      midiNote.getVelocity());
    context.write(REGION_IMAGENO, NOTE_VELOCITY);
    regionCounter++;
  }
}
\end{lstlisting}
\caption{A portion of our mapper that takes color data to MIDI notes.
  For each pixel in the original image, the conversion in
  \code{Color2Music} is applied before setting the output key (as a
  region, number pair).}
\end{figure}
